# Reproducible Research: Peer Assessment 1

Sport Activity Analysis With Knitr
==================================

## Loading and preprocessing the data

Since the data has to be loaded from a zipped file, 
one would like to use unz().
This way the  html file can be attained by only running 
library(knitr) and knit2html("PA1_template.Rmd") from R.

```{r}
zipHandle<-unz("activity.zip", "activity.csv")
rawData<-read.table(zipHandle, header=TRUE, sep=",")
unlink(zipHandle)
rm(zipHandle)

```

As preprocessing the second column will be converted in a POSIXlt type.
Two other columns are also made in the right type 
and a dataframe is made from this.

```{r}
date<-as.POSIXlt(rawData$date)
steps<-as.integer(rawData$steps)
interval<-as.factor(rawData$interval)
dataFrame <- data.frame(date, interval, steps)
rm(rawData)
```

## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
